---
title: "HW1"
author: "Burak Şimşek - IE582 - Fall 2020"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE , warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, warning=FALSE ,message=FALSE}
#loading libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
```

#English Premier Leage Statistics from 2018-2019 season to 2020-2021 season

```{r read data, warning=FALSE ,message=FALSE}
#reading dataframes
data2020=read.csv("https://github.com/BU-IE-582/fall20-buraksimsekkk/raw/master/HW1/2020-2021.csv")
data2019=read.csv("https://github.com/BU-IE-582/fall20-buraksimsekkk/raw/master/HW1/2019-2020.csv")
data2018=read.csv("https://github.com/BU-IE-582/fall20-buraksimsekkk/raw/master/HW1/2018-2019.csv")

#combining datas
all_data <- bind_rows(data2020 , data2019 , data2018)
```


##Task 1
##Task 1.1

####First histogram shows the frequency of the goals of home team among 3 seasons.
####Second histogram shows the frequency of the goals of home team among 3 seasons.
####Third histogram shows the frequency of differance of the goals of home team and the goals of away team among 3 seasons.

```{r histograms}
#plotting histograms
hist(all_data$FTHG , breaks=10, col="blue" , xlab="Home Team Goals", ylab="Number of Games", main="Home Team Goals Histogram" , xaxp=c(0,10,10) , yaxp=c(0,450,5)) 
hist(all_data$FTAG , breaks=10, col="blue" , xlab="Away Team Goals", ylab="Number of Games", main="Away Team Goals Histogram" , xaxp=c(0,10,10))
hist((all_data$FTHG - all_data$FTAG) , breaks=20, col="blue" , xlab="Home Team Goals - Away Team Goals", ylab="Number of Games", main=" Differance of Goals Histogram" , xaxp=c(-9,9,18) , yaxp=c(0,180,6) ) 
```

##Task 1.2


##Task 2

##Task 2.1
4 bookmakers are chosen and their home win, draw and away win probabilities; P(Home Win), P(Draw) and P(Away Win) respectively, will be calculated. 


```{r bookmakers probability calculations, warning=FALSE ,message=FALSE}


probs  <- all_data %>% mutate( Home_prob_BWH = 1 / BWH,
                               Draw_prob_BWD = 1 / BWD, 
                               Away_prob_BWA = 1 / BWA,
                    
                               Home_prob_IWH = 1 / IWH,
                               Draw_prob_IWD = 1 / IWD,
                               Away_prob_IWA = 1 / IWA,
                    
                               Home_prob_PSH = 1 / PSH,
                               Draw_prob_PSD = 1 / PSD,
                               Away_prob_PSA = 1 / PSA,

                               Home_prob_WHH = 1 / WHH,
                               Draw_prob_WHD = 1 / WHD,
                               Away_prob_WHA = 1 / WHA)


```

##Task 2.2

```{r normalization of probability, warning=FALSE ,message=FALSE}


normalized  <- probs %>% mutate( norm_BWH = Home_prob_BWH / (Home_prob_BWH + Draw_prob_BWD + Away_prob_BWA),
                                 norm_BWD = Draw_prob_BWD / (Home_prob_BWH + Draw_prob_BWD + Away_prob_BWA),
                                 norm_BWA = Away_prob_BWA / (Home_prob_BWH + Draw_prob_BWD + Away_prob_BWA), 
                                                             
                                 norm_IWH = Home_prob_IWH / (Home_prob_IWH + Draw_prob_IWD + Away_prob_IWA),
                                 norm_IWD = Draw_prob_IWD / (Home_prob_IWH + Draw_prob_IWD + Away_prob_IWA),
                                 norm_IWA = Away_prob_IWA / (Home_prob_IWH + Draw_prob_IWD + Away_prob_IWA),
                                                             
                                 norm_PSH = Home_prob_PSH / (Home_prob_PSH + Draw_prob_PSD + Away_prob_PSA),
                                 norm_PSD = Draw_prob_PSD / (Home_prob_PSH + Draw_prob_PSD + Away_prob_PSA),
                                 norm_PSA = Away_prob_PSA / (Home_prob_PSH + Draw_prob_PSD + Away_prob_PSA),
                                  
                                 norm_WHH = Home_prob_WHH / (Home_prob_WHH + Draw_prob_WHD + Away_prob_WHA),
                                 norm_WHD = Draw_prob_WHD / (Home_prob_WHH + Draw_prob_WHD + Away_prob_WHA),
                                 norm_WHA = Away_prob_WHA / (Home_prob_WHH + Draw_prob_WHD + Away_prob_WHA)  )      

head(normalized, 5)
                                                      
```

##Task 2.3

